---
title: "BREAST CANCER WISCONSIN DATA LOGISTIC REGRESSION ANALYSIS "
author: "Hasan Ali ÖZKAN"
date: "10/30/2021"
output: html_document
---

#### Data downloaded from [here](https://www.kaggle.com/uciml/breast-cancer-wisconsin-data) (breast cancer wisconsin data-set).

##### First step we should include some library to use in our analysis.

```{r,echo=TRUE,include=TRUE,message=FALSE}
library(tidyverse)
library(knitr)
library(caret)
library(Amelia)
```

##### Let's read the data-set.

```{r,echo=TRUE,include=TRUE,message=FALSE,warning=FALSE}
data <-  as.data.frame(read_csv("breast_cancer_data.csv"))

```

##### Let's see is there any null values in our data-set? 

```{r,echo=TRUE,include=TRUE}
missmap(data)
```

##### **OPPS** We have an empty column let's remove it from the data.
```{r,echo=TRUE,include=TRUE}
data <- data[c(0:32)]
```

##### Let's look at our dataset.In order to see comfortably I took to transpose of the table. (Now our columns are rows and rows are columns in below.)  

```{r, include=TRUE,echo=TRUE}
data_to_show <- t(head(data))
kable(data_to_show,format="markdown")
```

##### Then let's look at our data types whether all of them are correct or not? By using ***str*** (structure) function. 
```{r, include=TRUE,echo=TRUE}
str(data)
```

##### As you see above we have a factor variable which is ***diagnosis*** (our dependent variable) but R assumes it is a character. Let's modify it to factor.

```{r,echo=TRUE,include=TRUE}
data$diagnosis <- as.factor(data$diagnosis)
```
##### Let's check to see if it's fixed.

```{r, include=TRUE,echo=TRUE}
is.factor(data$diagnosis)
```
##### Yeap we modify the ***diagnosis*** variable correctly.



##### Now it's time to look at the correlation between our columns.
```{r, include=TRUE,echo=TRUE}
correlations <- cor(data[c(3:32)])
corrplot::corrplot(correlations,method = "square",tl.cex = 0.6, tl.col = "black")
```

##### As you see above some of our columns have a strong positive correlation between them.That problem name is ***multicolinearity***. We don't want to see this but before handle this problems let's create a basic logistic regression model.

##### Firstly, we set a seed value to keep randomness always the same.
```{r, include=TRUE,echo=TRUE}

set.seed(42)
```

##### Then we split our data into train and test data. 
```{r, include=TRUE,echo=TRUE}
sample<- createDataPartition(y= data$diagnosis,p=0.8,list = FALSE)

train_data <- data[sample,]
test_data <- data[-sample,]
```

##### Now we are creating our model.
```{r, include=TRUE,echo=TRUE,warning=TRUE}
logistic_model <- glm(diagnosis~.,data = train_data,family = "binomial")

```
##### As you see above we get two warning messages. I researched these issues and found their explanation. [Here ](https://statisticsglobe.com/r-glm-fit-warning-algorithm-not-converge-probabilities), the author explains why these warning messages occur. I will quote a small part: 
> This depends heavily on the structure of the specific data set you are using. However, I advise to have a look at the pairwise correlations of the variables in your data to see if some of these correlations are unnaturally high.
It often happens that variables are included to a model by accident. Try to identify these variables and run your model again without these variables until the warning message does not appear anymore.


##### Let's see what it can be able to our model?

```{r,echo= TRUE,include=TRUE}
prob <- predict(logistic_model,newdata=test_data,type="response")
pred <- ifelse(prob > 0.5,"M","B")


confusionMatrix(test_data$diagnosis,as.factor(pred))


```
##### As you see above model accuracy is 96 %. It is avesome but we can increase this ratio by applying PCA(Principal Component Analysis) to our dataset. 

##### Let's do this in one line :)
```{r, include=TRUE,echo=TRUE}
pca <- prcomp(train_data[3:32], center = TRUE, scale = TRUE)

summary(pca)

```
##### When we look at the summary of our pca results the first 6 components enough for us since they are capture about 88.85% variability in the data.

##### Let's look at the result in a plot.
```{r, include=TRUE,echo=TRUE}
screeplot(pca, type = "l", npcs = 15, main = "Screeplot of the first 15 PCs")
abline(h = 1, col="red", lty=5)
legend("topright", legend=c("Eigenvalue = 1"),
       col=c("red"), lty=5, cex=0.6)
```

##### As you see above we have a line which name is ***Eigenvalue*** we can say ***eigenvalue*** is the standart deviation of our PC's and we just take the first 6 PC's which their eigenvalues greater than 1.

##### As in the below graph we can say again the first 6 components are capture about 88.85% variability in the data.
```{r,include=TRUE,echo=TRUE}
cumpro <- cumsum(pca$sdev^2 / sum(pca$sdev^2))
plot(cumpro[0:15], xlab = "PC #", ylab = "Amount of explained variance", main = "Cumulative variance plot")
abline(v = 6, col="red", lty=5)
abline(h = 0.88759, col="red", lty=5)
legend("topleft", legend=c("Cut-off @ PC6"),
       col=c("red"), lty=5, cex=0.6)

```

##### Now lets look at correlation again.
```{r,echo=TRUE,include=TRUE}
correlations <- cor(pca$x[,c(0:6)])

corrplot::corrplot(correlations,method = "square", tl.col = "black")

```

##### As you see above there is no correlation between our PC's.That's what we have wanted.

```{r,echo=TRUE,include=TRUE}
set.seed(42)

data_pca <- data.frame(diagnosis=train_data[,"diagnosis"],pca$x[,0:6])
head(data_pca)
```
##### Now we can create a more powerful model by using fewer columns.
```{r,echo=TRUE,include=TRUE,warning=FALSE}
set.seed(42)
model_pca <- glm(diagnosis ~ .,data= data_pca,family = binomial)

test_data_pca <- predict(pca,newdata = test_data)

```

##### Let's look at our model's performance metrics.
```{r,echo=TRUE,include=TRUE}

prob <- predict(model_pca , newdata = data.frame(test_data_pca[,1:6]),type = "response")

pred <- factor(ifelse(prob>0.5,"M","B"))

confusionMatrix(test_data$diagnosis,as.factor(pred))

```


##### As you see above our accuracy become 100 %.